---
title: "GapMinder"
author: "andreafernandes"
date: "29 March 2016"
output: html_document
---


******
####Billionaire Database

This database consists of the average age of Billionaires in the world in a four year time frame. Data is sourced from Forbes list of the World's billionaires in 2007. 

******

**Question to answer: Are the billionaire's of the world getting younger?**

******

```{r Chunk A, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
rm(list=ls())
gc()

setwd("/Users/andreafernandes/Desktop/Google Drive/Springboard_Data_Science_Course_2016/DataScience/Lesson3_EDA/3_GapMinder/")

Billionaire <- read.csv("/Users/andreafernandes/Desktop/Google Drive/Springboard_Data_Science_Course_2016/DataScience/Lesson3_EDA/3_GapMinder/IndicatorAverage.csv")

library(ggplot2)
library(dplyr)
```


Tidying data
      
        i) Renaming variables
        ii) Selecting relevant variables
        iii) Creating an Age_Year and age variable. 
              - Age_Year is the Year that Forbes released Billionaire age. 
              - age is Age of Billionaire at the time of release. 

```{r Chunk B,results="hide", message=FALSE}
head(data.frame(Billionaire))

table(is.na(Billionaire$X))

bdatabase <- Billionaire %>% 
                rename(Country = Average.age.of.billionaires, Age_Year2004 = X2004, Age_Year2005 = X2005, Age_Year2006 = X2006, Age_Year2007 = X2007) %>%
                select (Age_Year2004, Age_Year2005, Age_Year2006, Age_Year2007, Country)
```

```{r Chunk C, echo=TRUE, results="hide", message=FALSE, warning=FALSE}
library("tidyr", lib.loc="/usr/local/lib/R/3.2/site-library")
bd <- gather(bdatabase, Age_Year, age, Age_Year2004:Age_Year2007, na.rm = FALSE) 
```

******

*Plotting billionaire age distribution by each year provided by Forbes*

```{r Chunk C1, echo=FALSE, message=FALSE, comment=FALSE, warning=FALSE}
qplot(x = age, data = bd,
      binwidth = 1,
     # geom = "freqpoly",
      color = I('red'), 
      xlab = "Age Distribution") +
      scale_x_continuous(lim = c(20, 100), breaks = c(20, 100, 10)) +
      facet_wrap(~Age_Year)



qplot(x = age, data = bd,
      binwidth = 1,
      geom = "freqpoly",
      color = Age_Year, 
      xlab = "Age Distribution") +
      scale_x_continuous(lim = c(20, 100), breaks = c(20, 100, 10)) 

#it's a messy polygram(?) but you can see a little spike in at a yonger age in 2006 compared to 2004...
```

Compared to 2004, there seems to be more younger billionaires in 2006 and 2007.

*****

*Numerical Summaries*
```{r Chunk D, message=FALSE}
#need to get rid of the zero's to get sensible summaries
bd$age[bd$age==0] <- NA 

by(bd$age, bd$Age_Year, summary)
```
The numerical summary shows that the minimum age in 2007 is lower compared to the minimum age in 2004. 

******

**Question: Are the billionaire's of the world getting younger?**  
  
***Answer: There seems to be younger billionaires in 2007 compared to 2004.***