---
title: "Lesson4_Notes"
author: "andreafernandes"
date: "27 March 2016"
output: html_document
---
####Miscellaneous code and code to do exercises in Lesson 4. 

******


Loading data and setting up
```{r, Chunk A, warning=FALSE, comment=FALSE, message=FALSE}
rm(list=ls())
gc()

setwd("/Users/andreafernandes/Desktop/NewTestFolder/Lesson4_EDA/")
pf <- read.table("/Users/andreafernandes/Desktop/NewTestFolder/Lesson4_EDA/pseudo_facebook.tsv", sep="\t", header = TRUE)

library(ggplot2)
#names(pf)
```



Coord Trans
```{r Chunk B}
ggplot(aes(x =  age, y = friend_count), data = pf) + 
           geom_point(alpha = 1/20) +
           xlim(13,90) + 
           coord_trans( y = "sqrt")
```


Playing around with jitter and geompoint
```{r Chunk C}
ggplot(aes(x =  age, y = friendships_initiated), data = pf, title(main="Geompoint")) +
      geom_point()

ggplot(aes(x =  age, y = friendships_initiated), data = pf, title(main="geomjitter")) +
      geom_jitter()

```
The main difference between these two seems to be point is more rigid (there are more straight lines in this scatter plot) than jitter. Jitter may be more sensitive to the continuous age variable, whereas geompoint is more rigid. 


Sticking with geom_jitter as much as possible

******


Examine the relationship between friendships_initiated (y) and age (x) using the ggplot syntax.
```{r Chunk D}
ggplot(data = pf, aes(x =  age, y = friendships_initiated)) +
      geom_jitter(alpha = 1/15, 
                  position = position_nudge(), 
                  colour = I('red'), fill = I('#FF0000')) + 
      #xlim(13,90) +
      coord_trans(y = "sqrt") +
      scale_x_continuous(lim = c(10, 100), breaks = seq(10, 100, 10))
```

The younger the user, the more friend requests. The older the user, the fewer friend requests. 



******


Plot mean friend count vs. age using a line graph. Be sure you use the correct variable names and the correct data frame. You should be working with the new data frame created from the dplyr functions. The data frame is called 'pf.fc_by_age'.



Use geom_line() rather than geom_point to create the plot. You can look up the documentation for
geom_line() to see what it does.
```{r Chunk E}
library(dplyr)
pf.fc_by_age <-  pf %>%
                        group_by(age) %>%
                        summarise(friend_count_mean = mean(friend_count),
                                  friend_count_median = mean(friend_count), 
                                  n = n()) %>%
                        arrange(age)

head(pf.fc_by_age)


ggplot(data = pf.fc_by_age, aes(x =  age, y = friend_count_mean)) +
      geom_line(colour = 'red') + 
      xlim(13,90) +
      #coord_trans(y = "sqrt") +
      scale_x_continuous(lim = c(10, 100), breaks = seq(10, 100, 10))
```
The mean of the friend count is high for younger users, coompared to middle aged users. It spikes at 69 and mean friend counts increases at ridiculous older ages >=80 years to over 100 years.


******

Overlaying mean line over scatter plot, with upper and lower quantilies and median line
```{r Chunk F}
ggplot(data = pf, aes(x =  age, y = friend_count)) +
      geom_point(alpha = 1/20, 
                  #position = position_nudge(), 
                  colour = I('red'), 
                  fill = I('#FF0000')) + 
      xlim(13,90) +
      coord_trans(y = "sqrt") +
      geom_line(stat = 'summary', fun.y =  mean) +
      geom_line(stat = 'summary', fun.y = quantile, fun.args=list(probs = 0.25), linetype = 2, color = 'blue') +
      geom_line(stat = 'summary', fun.y = quantile, fun.args=list(probs = 0.5), linetype = 2, color = 'green') + 
      geom_line(stat = 'summary', fun.y = quantile, fun.args=list(probs = 0.75), linetype = 2, color = 'yellow') 
#+
#      coord_cartesian(xlim = c(67,70), ylim=c(0,500), expand=TRUE)
```


******

Correlation Exercise
```{r Chunk G}
cor.test(x = pf$age, y = pf$friend_count, method = "pearson")


#-0.02 - no meaningful correlation
# rule of thumb 

#> 0.3 or < - 0.3 meaningful correlation
# ~ 0.5 moderate correlation
# > 0.7 is large correlation
```


Create a scatterplot of likes_received (y) vs. www_likes_received (x). Use any of the
techniques that you've learned so far to modify the plot.

```{r Chunk H}
ggplot(data = pf, aes(x =  likes_received, y = www_likes_received)) +
      geom_point(alpha = 1/10, 
                  position = position_nudge(), 
                  colour = I('red'), fill = I('#FF0000')) +
      scale_x_continuous(lim = c(0, 5000), breaks = seq(0, 5000, 500)) +
      scale_y_continuous(lim = c(0, 3000), breaks = seq(0, 3000, 1000)) +
      #coord_cartesian(xlim = c(0, 100), ylim = c(0, 100)) +
      geom_smooth(method = 'lm', color = 'yellow')


#you could also do the above like this:

# ggplot(data = pf, aes(x =  www_likes_received, y = likes_received)) +
#      geom_point() +
#      xlim(0, quantile(pf$www_likes_received, 0.95)) +
#      ylim(0, quantile(pf$likes_received, 0.95)) +
#      geom_smooth(method = 'lm', color = 'red')
```
Summary: From the plot it looks like of all the likes received, the more the total number of likes the more they are from www. 

```{r Chunk I}
cor.test(x = pf$likes_received, y = pf$www_likes_received, method = "pearson", conf.level = 0.95)
# cor = 0.95 (0.947, 0.948)
```
There is a high correlation (not association, it's just a correlation) between the total likes and likes from www




******

**New Exercise**
This can't load, i went ahead used discussion to get a grasp on what the data wrangling others had done. 
```{r Chunk J}
install.packages('alr3')
library(alr3)
data(Mitchell)
```



******


Create a new variable, 'age_with_months', in the 'pf' data frame.
Be sure to save the variable in the data frame rather than creating
a separate, stand-alone variable. 

You will need to use the variables 'age' and 'dob_month' to create the variable 'age_with_months'.

Assume the reference date for calculating age is December 31, 2013.
```{r Chunk K}
str(pf)
pf$age_with_months <- pf$age + (1-((pf$dob_month)/12))
```

```{r Chunk L}
pf.fc_by_age_months <- pf %>% 
                              group_by(age_with_months) %>% 
                              summarise(friend_count_mean = mean(friend_count),
                                        friend_count_median = median(friend_count),
                                        n = n()) %>% 
                              arrange(age_with_months)

```



Create a new scatterplot showing friend_count_mean versus the new variable, age_with_months. Be sure to use the correct data frame (the one you create in the last exercise) AND subset the data to investigate users with ages less than 71.

```{r Chunk M}
ggplot((data = subset(pf.fc_by_age_months, age_with_months < 71)), 
        aes(x = age_with_months, y =  friend_count_mean)) +
      geom_point(alpha = 1, 
                  colour = I('red'), fill = I('#FF0000')) +
      scale_x_continuous(lim = c(0, 80), breaks = seq(0, 80, 10)) +
      scale_y_continuous(lim = c(0, 500), breaks = seq(0, 500, 50)) +
      #coord_cartesian(xlim = c(0, 100), ylim = c(0, 100)) +
      geom_smooth(method = 'lm', color = 'yellow')

```

```

