---
title: "Lesson3_Notes"
author: "andreafernandes"
date: "27 March 2016"
output: html_document
---

******

####Miscellaneous Notes and Codes written to Paste on Udacity Exercises in Lesson 3 

Here's the link: 

https://www.udacity.com/course/viewer#!/c-ud651/l-755618712/e-814098658/m-814098660


Setting data and Loading pseudo-facebook data
```{r, Chunk A, warning=FALSE, message=FALSE, comment=FALSE}
rm(list=ls())
gc()

setwd("/Users/andreafernandes/Desktop/Google Drive/Springboard_Data_Science_Course_2016/DataScience/Lesson3_EDA/1_Lesson3_EDA/")

pf <- read.table("./pseudo_facebook.tsv", sep="\t", header = TRUE)

#names(pf)
```


******

Friend count by gender
```{r Chunk B, warning=FALSE, message=FALSE, comment=FALSE}
suppressPackageStartupMessages(library("ggplot2", lib.loc="/usr/local/lib/R/3.2/site-library"))

qplot(x = friend_count, data = pf, binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) + 
  facet_wrap(~gender)

#ggsave("Friend_Count_by_Gender.png", last_plot())
```

Summary: 

These histograms are a count of fb users with friends ranging from 0 to 1000. The histogram shows more male fb users have 0 friends compared to women. More male fb users have 1 friend comapred to female fb users. It is still hard to interpret the histograms and determine which gender has more friends.

There is also an NA gender category, which needs to be appropriately excluded from analysis. 


To give a clearer picture of which gender has more fb friends, numerical summaries can be used, such as below:

```{r Chunk C, warning=FALSE, message=FALSE, comment=FALSE}
by(pf$friend_count, pf$gender, summary)
```

******

Duration on FB (tenure)
```{r Chunk D, warning=FALSE, message=FALSE, comment=FALSE}
qplot(x = tenure/365.25, data = pf, binwidth = 0.25,  color = I('white'), fill = I('#FF0000')) + 
      scale_x_continuous(lim = c(0, 7.5), breaks = seq(0, 8, 1))

#ggsave("Duration_on_FB.png", last_plot())
```

Summary: 

There are more people with shorter duration as members of fb than those with longer tenure. 

******

The above histogram (count of duration on fb) can be further explored by age...

But first is a histogram of fb users by age
```{r Chunk E, warning=FALSE, message=FALSE, comment=FALSE}
qplot(x = age, data = pf, 
      xlab = "Age of Users", 
      binwidth = 1, 
      color = I('white'), fill = I('#FF0000')) + 
  scale_x_continuous(lim = c(10, 100), breaks = seq(10, 100, 5))


#ggsave("AgeofFBusers.png", last_plot())
```

Summary: 

There are more younger aged fb user than older aged fb users. 

******

Tenure by age
```{r Chunk E1}
ggplot(aes(x = age, y = tenure/365.25), data = pf,  color = I('white'), fill = I('#FF0000')) + 
      geom_jitter(alpha =  1/20) +
      scale_y_continuous(lim = c(0, 7.5), breaks = seq(0, 8, 1))
```

Summary:

This scatter plot shows that there are many more younger age groups (13 to ~30 years) who have been a fb member for 1 to 2 years compared to older age groups. 

There are an unusally higher number of people who are above 90 years old who have a been a fb user from 0 to 7 years. This could be anonmalies or bad extreme data. 

There seems to be a consistent spread of users who have been members of fb for over 3 years. 


******

Arranging grid next to each other: USEFUL code!
```{r Chunk F, warning=FALSE, message=FALSE, comment=FALSE}
library("gridExtra", lib.loc="/usr/local/lib/R/3.2/site-library")
library("grid", lib.loc="/usr/local/Cellar/r/3.2.2_1/R.framework/Versions/3.2/Resources/library")

table(pf$friend_count)
# define individual plots
p1 <- qplot(x = friend_count, data = pf, binwidth = 50, color = I('white'), fill = I('#FF0000')) 

p2 <- qplot(x = log10(friend_count + 1), data = pf, binwidth = 0.1, color = I('white'), fill = I('#CC8C3C')) 

p3 <- qplot(x = sqrt(friend_count), data = pf, binwidth = 1, color = I('white'), fill = I('#F79421'))

# arrange plots in grid
grid.arrange(p1, p2, p3, ncol=1)


p4 <- qplot(x = friend_count, data = pf, binwidth = 50, color = I('white'), fill = I('#FF0000')) 

p5 <- qplot(x = friend_count, data = pf, binwidth = 0.1, color = I('white'), fill = I('#CC8C3C'))  + scale_x_log10()

p6 <- qplot(x = friend_count, data = pf, binwidth = 1, color = I('white'), fill = I('#F79421'))  + scale_x_sqrt()
grid.arrange(p4, p5, p6, ncol=1)

#ggsave("GridArrange.png", last_plot())
```


******

**Transforming the distribution to normalise it**

```{r Chunk G, warning=FALSE, message=FALSE, comment=FALSE}
#frequency polygon

#str(pf)

by(pf$www_likes, pf$gender, summary)

```
It is clear from the summary that this is a non normal distribution with a long tail (Highly "over dispersed" data). 

The summary also shows that for both genders, 50% of users give no www_likes. 

******

Number of Likes by Gender

Demo of the original plot, with no changes.
```{r Chunk G1, warning=FALSE, message=FALSE, comment=FALSE}
p7 <- qplot(x = www_likes, data = subset(pf, !is.na(gender)), 
            binwidth = 5, 
            color = I('white'), fill = I('#FF0000')) + 
       scale_x_continuous(name = "www_likes: Original With 0 www_likes", 
                          lim = c(0, 400), 
                          breaks = seq(0, 400, 50)) +
       scale_y_continuous(lim = c(0,4000), breaks = seq(0, 4000, 100)) +
      facet_wrap(~gender)

#ggsave("www_likes_all.png", last_plot())
```

Summary: 

Male users get fewer likes than women. (i.e. there are more men with 0 likes than women)

You can also see this in the numerical summary:

```{r Chunk G1.1}
by(pf$www_likes, pf$gender, summary)
```

Summary: 

From the plot and the numerical summary, the distribution of likes is not normal (more people have 0 likes than not). 

When doing regression analysis, we are assuming our distributions will be normal. In order to use non-normal distribution, we can TRANSFORM them into a normal distribution by taking the log of the x variable or the square root of the x variable (there are other ways, but these are main)  --- asked Guy question on this.

******

Transforming the original plot gives a normal - ish distribution.
```{r Chunk I, warning=FALSE, message=FALSE, comment=FALSE}
p9 <- qplot(x = www_likes, data = na.omit(pf), 
            binwidth = 0.1, 
            color = I('white'), fill = I('#FF0000'), 
            xlab="www_likes: Log10 Transformed") +
      facet_wrap(~gender) +
      scale_x_log10()

ggsave("www_likes_without_0likes_transformed.png", last_plot())
```


Just checking the sqrt plot (it's not as normal as log)
```{r Chunk J, warning=FALSE, message=FALSE, comment=FALSE}
p10 <- qplot(x = www_likes, data = na.omit(pf), 
             binwidth = 4, 
             color = I('white'), fill = I('#FF0000'), 
             xlab = "www_likes: Square Root Transformed") +
      facet_wrap(~gender) +
      scale_x_sqrt()
```

You have to think about this a bit more and decide what type of transofmation you would like to do. 

```{r Chunk K, warning=FALSE, message=FALSE, comment=FALSE}
grid.arrange(p7, p9, p10, ncol=1)

ggsave("www_likes_vs_gender.png", last_plot())
```


**Conclusion for exercise purposes:** Use the log10 transformed graph

******

**Frequency Polygon to overlay graphs**

You can **overlay** two graphs using teh "freqpoly" geom function.


Overlaying female histogram of www_likes over the male histogram of www_likes. 

```{r Chunk L, warning=FALSE, message=FALSE, comment=FALSE}
qplot(x = www_likes, data = na.omit(pf), 
      binwidth = 0.1, 
      geom = 'freqpoly', 
      color = gender, 
      xlab="www_likes: Log10 Transformed") +
      scale_x_log10()
```


Just checking what the SqRt Tranformation Gender Overlay would look like. You can use this for neater looking graphs. 

```{r Chunk M, warning=FALSE, message=FALSE, comment=FALSE}
qplot(x = www_likes, data = na.omit(pf), 
    binwidth = 4, 
    geom = 'freqpoly', 
    color = gender, 
    xlab = "www_likes: Square Root Transformed") +
      scale_x_sqrt()
```

Summary:

Compared to males, females have more www_likes; more males have less likes. This still cannot tell us whether OVERALL which gender has more likes. 

******

To interpret the freqpoly graph, you can use a numerical summary and add up number of fb www_likes by users, by gender 
```{r Chunk O, warning=FALSE, message=FALSE, comment=FALSE}
#key function
by(pf$www_likes, pf$gender, sum)

```
Summary:

Females have a lot more likes than males. 
******

**Doing Boxplots:** 

Friend count by gender
```{r Chunk P, warning=FALSE, message=FALSE, comment=FALSE}
qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot') + #, ylim = c(0,1000)) +
      scale_y_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 500))

#you could also use the ylim function or the coord cartesian layer..
#Coord cartesian matches the below summary stats exactly, as opposed to the scale_y_continuous or ylim.
```

Summary:

This plot shows us that more females have a slightly higher friend count than males. But it's not very clear.
Doing a numerical summary, can highlight the discrepancies in the boxplot further. 

```{r Chunk P1}
by(pf$friend_count, pf$gender, summary)
```



*******

**Coord cartesian to zoom**

Friend requests by gender
```{r Chunk Q, warning=FALSE, message=FALSE, comment=FALSE}
#str(pf)

#table(is.na(pf$friendships_initiated)) #no NA

by(pf$friendships_initiated, pf$gender, summary)
by(pf$friendships_initiated, pf$gender, sum)

qplot(x = gender, y = friendships_initiated, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot') + #, ylim = c(0,1000)) +
      coord_cartesian(ylim=c(0,150))

ggsave("FriendReq_byGender.png", last_plot())
```
