---
title: "Birthday_Exercise"
author: "andreafernandes"
date: "30 March 2016"
output: html_document
---

####Birthday Task
Your task is to investigate the distribution of your friends' birth months and days.

Here some questions you could answer:

      1) How many people share your birthday? 
      2) Which month contains the most number of birthdays?
      3) How many birthdays are in each month?
      4) Which day of the year has the most number of birthdays?
      5) Do you have at least 365 friends that have birthdays on everyday of the year?
      
******

**Slight issue**: The dob provided are all set in the year 2014. But the data processed, sets them in 2016. I will fix this!


**Now fixed**
******

```{r Chunk A, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
rm(list=ls())
gc()

setwd("/Users/andreafernandes/Desktop/Google Drive/Springboard_Data_Science_Course_2016/DataScience/Lesson3_EDA/5_Birthdays/")

Birthdays <- read.csv("/Users/andreafernandes/Desktop/Google Drive/Springboard_Data_Science_Course_2016/DataScience/Lesson3_EDA/5_Birthdays/Birthdays.csv")

library(ggplot2)
library(dplyr)
library("lubridate", lib.loc="/usr/local/lib/R/3.2/site-library")
```


******


Formatting dates: Chose to obtain at least the month and date. 

The year is being weird:

      "dateofbirth" will remain show the proper year (i.e.2014) if the year in "dates"" is "mm/dd/14", otherwise it is "2020"" if year in "dates" is 2014.... This has something to do with origin date. 
      
      I tried something to make it right..
      
      
```{r Chunk B, results = "hide"}
head(data.frame(Birthdays))
str(Birthdays)
bd <- Birthdays

bd <- Birthdays %>% mutate(dateofbirth = as.Date(dates, format = "%m/%d")) %>%
      mutate (dateofbirth = as.Date(dateofbirth, format = "%d/%m/%y")-(365.25*2.00011)) 
#to make sure that all "dateofbirth" value are in the year 2014 and to deal with the origin data adding years to the variable, hence "-(365.25*2.00011)"
```

```{r Chunk B1}
head(bd)
```
It looks like it worked. 

******

To plot date variables as a histogram, there are a few cuts that need to be made. I've tried cut date by month, week and day by adding these variables to the Birthday (bd) dataset.

```{r Chunk C, message=FALSE, warning=FALSE}
bd$Month <- as.Date(cut(bd$dateofbirth, breaks = "month"))
bd$Week <- as.Date(cut(bd$dateofbirth, breaks = "week", start.on.monday = FALSE)) # changes weekly break point to Sunday
bd$Day <- as.Date(cut(bd$dateofbirth, breaks = "day"))
```           


I can then use scale_x_date to break histogram by weeks, month or year and label by them too. 
```{r Chunk C2}
qplot(x = dateofbirth, color = I('white'), fill = Month, data = bd, 
      binwidth = 1) +
      facet_wrap(~Month, scales = "free_x")
```
Will need to play aorund with this a bit more. 

I can now use it and numerical summaries to answer question on this database


******


1) How many people share your birthday?
```{r Chunk D}
library(dplyr)

bd %>% 
        filter(month(dateofbirth) == 8 & day(dateofbirth) == 25) %>%
        group_by(month(dateofbirth), day(dateofbirth)) %>% summarise(count = n()) %>% data.frame()

# Three birthdays on the 25th of August, 2014
```


2) Which month contains the most number of birthdays?
```{r Chunk E}
bd %>% group_by(month(dateofbirth)) %>% summarise(count = n()) %>% arrange(count) #March
```


3) How many birthdays are in each month? 
```{r Chunk F} 
bd %>% group_by(month(dateofbirth)) %>% summarise(count = n())
```


4) Which day of the year has the most number of birthdays? 


```{r Chunk G1}
qplot(x = dateofbirth, color = I('white'), fill = Month, data = bd, 
      binwidth = 1) +
      facet_wrap(~Month, scales = "free_x")

# 5th Feb, 22nd May and 16 July
```



5) Do you have at least 365 friends that have birthdays on everyday of the year?
```{r Chunk H}
#See Plot -  No
```



******


      
      