---
title: "EDA Problem Set 5: EDA"
author: "andreafernandes"
date: "9 April 2016"
output: 
  html_document:
    keep_md: true
    theme: united
    highlight: tango
    toc: true
    toc_depth: 3
---

******

###Introduction and Background for Gapminder Problem Set 5 Exercise

```{r Chunk A}
# https://www.udacity.com/course/viewer#!/c-ud651/l-770038733/e-867869590/m-901058638

# The Gapminder website contains over 500 data sets with information about
# the world's population. Your task is to continue the investigation you did at the
# end of Problem Set 4 or you can start fresh and choose a different
# data set from Gapminder.

# If you’re feeling adventurous or want to try some data munging see if you can
# find a data set or scrape one from the web.

# In your investigation, examine 3 or more variables and create 2-5 plots that make
# use of the techniques from Lesson 5.

# You can find a link to the Gapminder website in the Instructor Notes.

# Once you've completed your investigation, create a post in the discussions that includes:
#       1. the variable(s) you investigated, your observations, and any summary statistics
#       2. snippets of code that created the plots
#       3. links to the images of your plots

# Copy and paste all of the code that you used for
# your investigation, and submit it when you are ready.
# ============================================================================================
```

```{r Chunk A1, echo = FALSE, results= "hide"}
rm(list=ls())
gc()


# for toc
#http://rmarkdown.rstudio.com/html_document_format.html

suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(rmarkdown))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(markdown))
suppressPackageStartupMessages(library(stargazer))
suppressPackageStartupMessages(library(gmodels))
```

```{r Chunk A1.1, echo = FALSE, results= "hide"}
setwd("/Users/andreafernandes/Desktop/Google Drive/Springboard_Data_Science_Course_2016/DataScience/Lesson5_EDA/GapMinder_Project/Lesson5_EDA/Gapminder_project_Antidepressants/")

ad <- read.csv("/Users/andreafernandes/Desktop/Google Drive/Springboard_Data_Science_Course_2016/DataScience/Lesson5_EDA/GapMinder_Project/Lesson5_EDA/Gapminder_project_Antidepressants/Antidepressant_Profiling_2014_Cohort_v2.csv", header = TRUE)
```


```{r Chunk A2, echo = FALSE, results= "hide"}
# Cleaning the data

#|**TrazoStartDate**| **INCORRECT**: rename this variable TO TrazoFlag. 0/1 binary variable indicating whether patient was on this drug during snapshot

# |**MirtStartDate**| **INCORRECT** should be converted into a binary variable, 1 or 0. and renamed MirtFlag |

ad <- ad %>% mutate(MirtStartDate = as.Date(MirtStartDate, format = "%d/%m/%Y")) %>%
             mutate(MirtazapineFlag = ifelse(year(MirtStartDate) < 1901, 0, 1)) %>%
             rename(OtherAD_TrazodoneFlag = TrazoStartDate)

#str(ad)

# test out date formats with this code: as.Date(ad$MirtStartDate[1:20], format="%d/%m/%Y")

# compare old and new columns with this code: ad[,c("MirtFlag","MirtStartDate")]

# lots of work in frormating dates! This works `as.Date(MirtStartDate, format = "%d/%m/%Y")
```



******

###My "Gapminder" project Aims and Objectives using the Antidepressant Dataset

####**Part 1: Antidepressant Use in Secondary Mental Health Care Patients** 

The aim of this part of the project is to understand how antidepressants are used among depressed patients in this secondary mental health care setting. In other words, what are the common prescriptions of antidepressants in this secondary mental health care facility? How are antidepressants prescribed? Is the entire cohort receiving antidepressant treatment?

####**Part 2: Comparing Antidepressant users to Non antidepressant users**

All patients in this cohort have a diagnosis of depression. About half the individuals in this cohort are not on any antidepressants during the snapshot. Are there any differences in the patients using antidepressants compared to patients not using antidepressants?

####**Part 3: Antidepressant Use and Suicide Attempt**

There is some indication that one of the side-effects of antidepressant use is suicidality (i.e. suicidal ideation, suicdie attempt or death by suicide) [ref]. Conducting preliminary analysis, is there any indication that use of antidepressant may be associated with suicide attempts? 

******

******
###THE DATASET

**Description of the Antidepressant dataset.**

The dataset used here is a 6 month snapshot of antidepressant use among mental health patients with a clinical diagnosis of depression. 

**Data Source**

The data is derived from a mental health clinical trust based in South London and covers an area with a population of around 1.4 million residents. The trust provides secondary mental health care to individuals who are referred by GPs, privately referred, A&E and self-referrals.

In 2008, a research facility was founded which used pseudonymised the electronic clinical records from the South London trust for research and clinical audit purposes. Currently there are ~250000 records within this research database. Within the database, there are around 70000 individuals with a clinical diagnosis of dperession, who were known to the trust (comorbid depression included). 

**The Sample**

To investigate or explore AntiDepressant (AD) use I selected individuals who met both of the following criteria:

|_Inclusion Criteria_|
|:-------|
| 1) Patient had a clinical diagnosis of dpression but no other major mental disorder (such as schizophrenia, bipolar disorder etc)|
| 2) Patient had some face to face contact with clinical staff at some point during the 1st of March 2014 to 31st August 2014. |

******

The sample consisted of 4157 patients. 

*******

*******

|_**Acronym Key**_|  |
|:-------|:-------|
|AD | AntiDepressant |
|Snapshot period | Randomly selected 6 month period from 1st of March, 2014 to the 31st of August, 2014|
|LSOA code | Lower Super Output Area code (it's an area level address code) |
|TCA | Antidepressant class: TriCyclic Antidepressants|
|MAOI | Antidepressant class: Mono Amine Oxidase Inhibitors |
|SSRI | Antidepressant class: Selective Serotonin Reuptake Inhibitors|
|SNRI | Antidepressant class: Selective Norepinephrine Reuptake Inhibitors|


******

```{r Chunk B, echo = FALSE, results= "hide"}

ad_1 <- ad %>% 
# selecting the relevant variables (i.e. IDs and drugs)
              select(a, 
                     PatientonAD, 
                     AMITRIPTYLINE = TCA_AmitriptylineFlag, 
                     CLOMIPRAMINE = TCA_CLOMIPRAMINEFlag, 
                     DOSULEPIN = TCA_DOSULEPINFlag, 
                     DOXEPIN = TCA_DOXEPINFlag, 
                     IMIPRAMINE = TCA_IMIPRAMINEFlag, 
                     LOFEPRAMINE = TCA_LOFEPRAMINEFlag, 
                     NORTRIPTYLINE = TCA_NORTRIPTYLINEFlag, 
                     TRIMIPRAMINE = TCA_TRIMIPRAMINEFlag, 
                     MIANSERIN = TCARELATED_MIANSERINFlag, 
                     ISOCARBOXAZID = MAOI_ISOCARBOXAZIDFlag, 
                     PHENELZINE = MAOI_PHENELZINEFlag, 
                     TRANYLCYPROMAINE = MAOI_TRANYLCYPROMAINEFlag, 
                     MOCLOBEMIDE = REVERSEMAOI_MOCLOBEMIDEFlag, 
                     CITALOPRAM = SSRI_CITALOPRAMFlag, 
                     ESCITALOPRAM = SSRI_ESCITALOPRAMFlag, 
                     FLUOXETINE = SSRI_FLUOXETINEFlag, 
                     FLUVOXAMINE = SSRI_FLUVOXAMINEFlag, 
                     PAROXETINE = SSRI_PAROXETINEFlag, 
                     SERTRALINE = SSRI_SERTRALINEFlag, 
                     DOLUXETINE = SNRI_DOLUXETINEFlag, 
                     REBOXETINE = SNRI_REBOXETINEFlag, 
                     VENLAFAXINE = OTHERAD_VENLAFAXINEFlag, 
                     AGOMELATIN = OTHERAD_AGOMELATINFlag, 
                     BUPROPION = OTHERAD_BUPROPIONFlag, 
                     MIRTAZAPINE = MirtazapineFlag, 
                     TRYPTOPHAN = OTHERAD_TRYPTOPHANFlag,
                     TRAZODONE = OtherAD_TrazodoneFlag) %>% 
#creating one drug class and name column and one Flag column
              gather(AD_Type_and_Name, Binary_AD_Flag, AMITRIPTYLINE:TRAZODONE, na.rm = FALSE)
```

******

******

###Part 1: ANTIDEPRESSANT USE
```{r Chuck C, echo=FALSE, warning =FALSE}
ggplot(aes(x = AD_Type_and_Name, y = Binary_AD_Flag, 
           color = AD_Type_and_Name), data =  ad_1) +
      geom_bar(stat = "identity") +
      theme(legend.position = "none", axis.text.x = element_text(angle = 60, hjust = 1)) +
      ggtitle("Use of Antidepressants") +
      ylab("Number of Patients on Antidepressants")
```

******

The plot above shows the number of patients within each antidepressant prescribed at the Trust. The tables below provide the corresponding numbers.

******

**Number of patients on each antidepressant.**
```{r Chunk C1, echo=FALSE}
ad_1 %>% filter(Binary_AD_Flag == 1) %>%
          group_by(AD_Type_and_Name) %>%
          summarise(count = n()) %>% 
          arrange(count) %>% 
          data.frame() %>% kable(align = "l")

# Excluding Other AD Tryptophan from this list because it is not at all used in this sample. 
# Clinician impression is to exclude Tryptophan entirely from the analysis.  
```

___What's the most common antidepressant prescribed?___

      - Mirtazapine, an atypical antidepressant with noradrenergic and specific serotonergic activity, is the 
      
        MOST COMMONLY prescribed / used antidepressant within the 6-month snapshot period. 
      
      - Followed by SSRI Sertraline, SSRI Citalopram, SNRI Venlafaxine and SSRI Fluoxetine.

******

```{r Chunk C2, echo= FALSE, results = "hide"}
# Make a variable that means "type" of AD user: Single AD, Single class and Multiple class AD user. 

# ggg <- ad_3 %>%  mutate(Number_of_Drugs = rowSums(ad_3[,3:29], na.rm = TRUE))

#ad_3 <- ad_2 %>% 
#              filter(PatientonAD == 1) 

## count number of drug for each sample 
## the easy way to calc rowSums on specific columns is to use R's [row, cols] indexing syntax as above
```


*****

####Type of Antidepressant Users

Are patients being prescribed single antidepressant drugs or in combination?

******

```{r Chunk J, echo=FALSE, results="hide"}
ad$SNRI_DOLUXETINEFlag <- as.numeric(ad$SNRI_DOLUXETINEFlag)
ad$SNRI_REBOXETINEFlag <- as.numeric(ad$SNRI_REBOXETINEFlag)
ad$OTHERAD_VENLAFAXINEFlag <- as.numeric(ad$OTHERAD_VENLAFAXINEFlag)
ad$TCA_AmitriptylineFlag <- as.numeric(ad$TCA_AmitriptylineFlag)
ad$TCA_CLOMIPRAMINEFlag <- as.numeric(ad$TCA_CLOMIPRAMINEFlag )
ad$TCA_DOSULEPINFlag <- as.numeric(ad$TCA_DOSULEPINFlag)
ad$TCA_DOXEPINFlag <- as.numeric(ad$TCA_DOXEPINFlag)
ad$TCA_IMIPRAMINEFlag <- as.numeric(ad$TCA_IMIPRAMINEFlag)
ad$TCA_LOFEPRAMINEFlag <- as.numeric(ad$TCA_LOFEPRAMINEFlag)
ad$TCA_NORTRIPTYLINEFlag <- as.numeric(ad$TCA_NORTRIPTYLINEFlag)
ad$TCA_TRIMIPRAMINEFlag <- as.numeric(ad$TCA_TRIMIPRAMINEFlag)
ad$TCARELATED_MIANSERINFlag <- as.numeric(ad$TCARELATED_MIANSERINFlag)
ad$MAOI_ISOCARBOXAZIDFlag <- as.numeric(ad$MAOI_ISOCARBOXAZIDFlag)
ad$MAOI_PHENELZINEFlag <- as.numeric(ad$MAOI_PHENELZINEFlag)
ad$MAOI_TRANYLCYPROMAINEFlag <- as.numeric(ad$MAOI_TRANYLCYPROMAINEFlag)
ad$REVERSEMAOI_MOCLOBEMIDEFlag <- as.numeric(ad$REVERSEMAOI_MOCLOBEMIDEFlag)
ad$SSRI_CITALOPRAMFlag <- as.numeric(ad$SSRI_CITALOPRAMFlag)
ad$SSRI_ESCITALOPRAMFlag <- as.numeric(ad$SSRI_ESCITALOPRAMFlag)
ad$SSRI_FLUOXETINEFlag <- as.numeric(ad$SSRI_FLUOXETINEFlag )
ad$SSRI_FLUVOXAMINEFlag <- as.numeric( ad$SSRI_FLUVOXAMINEFlag )
ad$SSRI_PAROXETINEFlag <- as.numeric(ad$SSRI_PAROXETINEFlag ) 
ad$SSRI_SERTRALINEFlag <- as.numeric(ad$SSRI_SERTRALINEFlag )
ad$OTHERAD_AGOMELATINFlag <- as.numeric(ad$OTHERAD_AGOMELATINFlag )
ad$OTHERAD_BUPROPIONFlag <- as.numeric(ad$OTHERAD_BUPROPIONFlag )
ad$OtherAD_TrazodoneFlag <- as.numeric(ad$OtherAD_TrazodoneFlag )
ad$MirtazapineFlag <- as.numeric(ad$MirtazapineFlag )
ad$LatestNumberOfSpells <- as.numeric(ad$LatestNumberOfSpells)
ad$Adjusted_Total <- as.numeric(ad$Adjusted_Total)

```


```{r Chunk J1, echo=FALSE, results="hide"}
ad_3 <- ad %>% select(a, PatientonAD,
         TCA_AmitriptylineFlag, 
         TCA_CLOMIPRAMINEFlag, 
         TCA_DOSULEPINFlag, 
         TCA_DOXEPINFlag, 
         TCA_IMIPRAMINEFlag, 
         TCA_LOFEPRAMINEFlag, 
         TCA_NORTRIPTYLINEFlag, 
         TCA_TRIMIPRAMINEFlag, 
         TCARELATED_MIANSERINFlag, 
         MAOI_ISOCARBOXAZIDFlag, 
         MAOI_PHENELZINEFlag, 
         MAOI_TRANYLCYPROMAINEFlag, 
         REVERSEMAOI_MOCLOBEMIDEFlag, 
         SSRI_CITALOPRAMFlag, 
         SSRI_ESCITALOPRAMFlag, 
         SSRI_FLUOXETINEFlag, 
         SSRI_FLUVOXAMINEFlag, 
         SSRI_PAROXETINEFlag, 
         SSRI_SERTRALINEFlag, 
         SNRI_DOLUXETINEFlag, 
         SNRI_REBOXETINEFlag, 
         OTHERAD_VENLAFAXINEFlag, 
         OTHERAD_AGOMELATINFlag, 
         OTHERAD_BUPROPIONFlag, 
         OtherAD_TrazodoneFlag, 
         MirtazapineFlag,
         primary_diagnosis,
         LithiumFlag,
         ethnicitycleaned,
         cleaneddateofbirth,
         Gender_ID,
         imd_score,
         Marital_Status_ID,
         LatestNumberOfSpells,
         Accepted_Date,
         Agitated_Behaviour_Score_ID,
         Self_Injury_Score_ID,
         Problem_Drinking_Drugs_Score_ID,
         Cognitive_Problems_Score_ID,
         Physical_Illness_Score_ID,
         Hallucinations_Score_ID,
         Depressed_Mood_Score_ID,
         Other_Mental_Problems_Score_ID,
         Other_Mental_Problems_Type_ID,
         Relationship_Problems_Score_ID,
         Daily_Living_Problems_Score_ID,
         Living_Conditions_Problems_Score_ID,
         Occupational_Problems_Score_ID,
         Adjusted_Total,
         PBR_Repeat_Self_Harm_ID,
         Rating_Date,
         noncompliant,
         compliant, 
         PTSessionDate,
         SessionNumber,
         AntiPsychoticDate,
         AntiPsychoticFlag, 
         PastADdate,
         PastAD, 
         BeenInpatientin12months,                                           
         Address_Start,
         RecentHallucinations,
         SixmonthsPastHallucinations,
         PoorMotivationPast,
         PoorAffectPast,  
         AnhedoniaPast,  
         HopelessnessPast,  
         WorthlessnessPast, 
         SuicideAttemptPast,
         AnergiaPast, 
         WorthlessnessRecent,
         HopelessnessRecent,
         AnhedoniaRecent,
         PoorAffectRecent,
         PoorMotivationRecent,
         SuicideAttemptRecent,
         AnergiaRecent)
```

```{r Chunk J2, echo=FALSE, results="hide"}
ad_3 <- ad_3 %>%
                 mutate(TCA.Total = rowSums(ad_3[,3:11], na.rm = TRUE)) %>%
                 mutate(MAOI.Total = rowSums(ad_3[,12:15], na.rm = TRUE)) %>%
                 mutate(SSRI.Total = rowSums(ad_3[,16:21], na.rm = TRUE)) %>%
                 mutate(SNRI.Total = rowSums(ad_3[,22:24], na.rm = TRUE)) %>%
                 mutate(OtherAD.Total = rowSums(ad_3[,25:27], na.rm = TRUE)) %>%
                 mutate(Mirt.Total = MirtazapineFlag, na.rm = TRUE) 
```

```{r Chunk J3, echo=FALSE, results="hide"} 

ad_3 <- ad_3 %>%
        mutate(Number_of_Drugs = rowSums(ad_3[,3:28], na.rm = TRUE)) %>%
        mutate(Type.of.user = ifelse(Number_of_Drugs %in% 1, "Single", 
                              ifelse(Number_of_Drugs %in% 0, "Not on Antidepressants", "Multiple"))) 
```

```{r Chunk J4, echo=FALSE, results="hide"}
ad_3 <- ad_3 %>% filter(Type.of.user == "Multiple" | Type.of.user == "Single") %>%
                 mutate(Type.of.multiple.AD.user = ifelse(Number_of_Drugs == TCA.Total, "TCA.class.AD.user",
                                                   ifelse(Number_of_Drugs == MAOI.Total, "MAOI.class.AD.user", 
                                                   ifelse(Number_of_Drugs == SSRI.Total, "SSRI.class.AD.user", 
                                                   ifelse(Number_of_Drugs == SNRI.Total, "SNRI.class.AD.user",
                                                   ifelse(Number_of_Drugs == Mirt.Total, "Mirt.class.AD.user",
                                                   ifelse((TCA.Total == 1 & MAOI.Total == 1), "TCA.MAOI",
                                                   ifelse((TCA.Total == 1 & SSRI.Total == 1), "TCA.SSRI",
                                                   ifelse((TCA.Total == 1 & SNRI.Total == 1), "TCA.SNRI",
                                                   ifelse((TCA.Total == 1 & OtherAD.Total == 1), "TCA.OtherAD",
                                                   ifelse((TCA.Total == 1 & Mirt.Total == 1), "TCA.Mirtazapine",
                                                   ifelse((MAOI.Total == 1 & SSRI.Total == 1), "MAOI.SSRI",
                                                   ifelse((MAOI.Total == 1 & SNRI.Total == 1), "MAOI.SNRI",
                                                   ifelse((MAOI.Total == 1 & OtherAD.Total == 1), "MAOI.OtherAD",
                                                   ifelse((MAOI.Total == 1 & Mirt.Total == 1), "MAOI.Mirtazapine",
                                                   ifelse((SSRI.Total == 1 & SNRI.Total == 1), "SSRI.SNRI",
                                                   ifelse((SSRI.Total == 1 & OtherAD.Total == 1), "SSRI.OtherAD",
                                                   ifelse((SSRI.Total == 1 & Mirt.Total == 1), "SSRI.Mirtazapine",
                                                   ifelse((SNRI.Total == 1 & OtherAD.Total == 1), "SNRI.OtherAD",
                                                   ifelse((SNRI.Total == 1 & Mirt.Total == 1), "SNRI.Mirtazapine", 
                                                   ifelse((OtherAD.Total == 1 & Mirt.Total == 1), "OtherAD.Mirtazapine", "ThreeAndMore.classes.AD.user")))))))))))))))))))))

# notes:
# If TCA.Total == Number_of_Drugs, then "Single Class Prescription"
#    MAOI.Total == Number_of_Drugs, then "Single Class Prescription"
#    SSRI.Total == Number_of_Drugs, then "Single Class Prescription"
#    SNRI.Total == Number_of_Drugs, then "Single Class Prescription"
#    OtherAD.Total == Number_of_Drugs, then "Single Class Prescription"
#    Mirt.Total == Number_of_Drugs, then "Single Class Prescription"
# If not, then "Multiple Class Prescription"

```

```{r Chunk J5, echo= FALSE, results = "hide"}
head(ad_3[,c("Number_of_Drugs","TCA.Total", "MAOI.Total", "SSRI.Total", "SNRI.Total", "OtherAD.Total", "Mirt.Total", "Type.of.multiple.AD.user")])
```

```{r Chunk J5.1, echo=FALSE}
qplot(Type.of.user, data = ad_3, fill = Type.of.user) + 
  labs(title = "Type of Antidepressant Users")
```

******

Antidepressants are most commonly prescribed as single antidepressants.

******

```{r Chunk J6, echo=FALSE}
qplot(Type.of.multiple.AD.user, data = ad_3, fill = Type.of.multiple.AD.user) +
      labs(title = "Type of Multiple Antidepressant User") +
      theme(legend.position = "none", axis.text.x = element_text(angle = 60, hjust = 1)) 
```

******

Patients who are prescribed multiple antidepressants get prescribed different classes of antidepressants as opposed to antidepressants within the same class (e.g. SSRI in combination with an SNRI as opposed to another SSRI).

*******

Here's corresponding numbers to the graphs above. 

####**Most Common Multiple Antidepressant user groups**               
```{r Chunk J7, echo=FALSE}
ad_3 %>%  group_by(Type.of.multiple.AD.user) %>%
          summarise(count = n()) %>% 
          arrange(count) %>% 
          data.frame() %>%
          kable(align = "l")
```

******

####**Most common Single antidepressant user groups**
```{r Chunk L, echo=FALSE}
ad_3 %>% filter(Type.of.user == "Single") %>% 
          gather(Antidepressant, Flag, TCA_AmitriptylineFlag:MirtazapineFlag) %>% 
          filter(Flag == 1) %>%
          group_by(Antidepressant) %>%
          summarise(count = n()) %>% 
          arrange(count) %>% 
          data.frame() %>% kable(align = "l")
```



```{r Chunk M, echo=FALSE, results="hide"}


#__The most common mutually exclusive "antidepressant user" groups__

#      In light of how antidepressants are being prescribed, any further analysis of antidepressants use will 
      
#      be made in three separate cohorts based on antidepressant usage:
      
#                  i) patients who are prescribed single antidepressants.
                  
#                  ii) patients who are prescribed multiple antidepressants from the SAME CLASS
                  
#                  iii) patients who are prescribed multiple antidepressants from different classes
                  
#                       (e.g. patients who are prescribed SSRI and SNRI drugs will be groups together)



#Db.i implies Database 1
Db.i <- filter(ad_3, Type.of.multiple.AD.user == "Single AD" | Type.of.multiple.AD.user == "Not on AD")

str(Db.i)
```

```{r Chunk N, echo=FALSE, results="hide"}
#Db.ii implies Database 2
Db.ii <- filter(ad_3, Type.of.multiple.AD.user == "SNRI.class.AD.user" | 
                      Type.of.multiple.AD.user == "SSRI.class.AD.user" | 
                      Type.of.multiple.AD.user == "TCA.class.AD.user" | 
                      Type.of.multiple.AD.user == "Not on AD")

str(Db.ii)
```                                       

```{r Chunk O, echo=FALSE, results="hide"}
#Db.iii implies Database 2
Db.iii <- filter(ad_3, Type.of.multiple.AD.user == "MAOI.SNRI" | 
                      Type.of.multiple.AD.user == "MAOI.SSRI" | 
                      Type.of.multiple.AD.user == "Multi.class.AD.user" | 
                      Type.of.multiple.AD.user == "OtherAD.Mirtazapine" | 
                      Type.of.multiple.AD.user == "SNRI.Mirtazapine" | 
                      Type.of.multiple.AD.user == "SNRI.OtherAD" | 
                      Type.of.multiple.AD.user == "SSRI.OtherAD" | 
                      Type.of.multiple.AD.user == "SSRI.SNRI" | 
                      Type.of.multiple.AD.user == "TCA.MAOI " | 
                      Type.of.multiple.AD.user == "TCA.Mirtazapine" | 
                      Type.of.multiple.AD.user == "TCA.SNRI" | 
                      Type.of.multiple.AD.user == "TCA.SSRI" | 
                      Type.of.multiple.AD.user == "Not on AD")

str(Db.iii)
```     

******

```{r Chunk D, echo = FALSE, results = "hide"}
# assigning 1 and 0 as values for Patient on AD and Patient off AD, respectivley, in the 6 months. 

ad$PatientonAD[is.na(ad$PatientonAD)] <- 0 # Not on AD in snapshot 6 months
ad$PatientonAD[ad$PatientonAD != 0] <- 1 # On AD in snapshot 6 months

# ad_2 %>% head(20)

#Number of Individuals on AD vs off AD
a <- ad %>% 
        group_by(PatientonAD) %>% 
        summarise(count = n(),
        percentage = (n()/sum(2039, 2118))*100)

a$PatientonAD[a$PatientonAD == 0] <- "No"
a$PatientonAD[a$PatientonAD == 1] <- "Yes"

#PatientonAD: 1 - Patient on AD during snapshot; 2 - Patient off AD during snapshot
```

___Are all patients with depression prescribed antidepressants?___


```{r Chunk E1, echo = FALSE}
kable(a, align = "c", caption = "Number of Patients on Antidepressants versus off Antidepressants", digits = 2)
```

There seems to be a 50:50 split, about half the patients are on AD and the other half not on antidepressants at the time of the study snapshot. Let's compare the two groups to see if they are different in any way...

```{r Chunk F1, echo = FALSE, results = "hide"}
# useful link on chi^2 test: http://www.stat.yale.edu/Courses/1997-98/101/chisq.htm

ad <- ad %>%
          mutate(dob = as.Date(cleaneddateofbirth, format = "%d/%m/%Y")) %>%
          mutate(age = year(as.Date("01/01/2014", format = "%d/%m/%Y")) - year(dob)) %>%
          mutate(Accepted_Date = as.Date(Accepted_Date, format = "%d/%m/%Y")) %>%
          mutate(time.at.mh.service = abs(date(as.Date("01/01/2014", format = "%d/%m/%Y")) - date(Accepted_Date))) %>%
          mutate(imd_quartiles = cut(imd_score, b = 4)) %>%
          mutate(Adjusted_Total_level = ifelse(Adjusted_Total < 25, "Mild",
                                        ifelse(Adjusted_Total >=25 & Adjusted_Total < 51, "Moderate",
                                        ifelse(Adjusted_Total >=51 & Adjusted_Total < 101, "Severe", "unknown")))) %>%
          mutate(ethnicity = ifelse(ethnicitycleaned == "African (N)", "Black",
                             ifelse(ethnicitycleaned == "Any other Asian background (L)", "Asian", 
                             ifelse(ethnicitycleaned == "Any other black background (P)", "Black",
                             ifelse(ethnicitycleaned == "Any other ethnic group (S)", "Other",
                             ifelse(ethnicitycleaned == "Any other mixed background (G)", "Other",
                             ifelse(ethnicitycleaned == "Any other white background (C)", "White",
                             ifelse(ethnicitycleaned == "Bangladeshi (K)", "Asian",
                             ifelse(ethnicitycleaned == "British (A)", "White",
                             ifelse(ethnicitycleaned == "Caribbean (M)", "Black",
                             ifelse(ethnicitycleaned == "Chinese (R)", "Asian",
                             ifelse(ethnicitycleaned == "Indian (H)", "Asian",
                             ifelse(ethnicitycleaned == "Irish (B)", "White",
                             ifelse(ethnicitycleaned == "Not Stated (Z)", "Other",
                             ifelse(ethnicitycleaned == "Pakistani (J)", "Asian",
                             ifelse(ethnicitycleaned == "White and Asian (F)", "Other",
                             ifelse(ethnicitycleaned == "White and Black African (E)", "Other",
                             ifelse(ethnicitycleaned == "White and black Caribbean (D)", "Other",
                             ifelse(ethnicitycleaned == "xNx", "Other", "Other"))))))))))))))))))) %>%
          mutate(marital_status = ifelse((Marital_Status_ID == "Cohabiting" | Marital_Status_ID == "Married" | Marital_Status_ID == "Married/Civil Partner"), "Married",
                                  ifelse((Marital_Status_ID == "Divorced" | Marital_Status_ID == "Divorced/Civil Partnership Dissolved" | Marital_Status_ID == "Separated"), "Separated",
                                  ifelse((Marital_Status_ID == "Widowed/Surviving Civil Partner" | Marital_Status_ID ==  "Widowed"), "Widowed",
                                  ifelse((Marital_Status_ID == "Not Disclosed" | Marital_Status_ID ==  "Not Known"), "Unknown",
                                  ifelse((Marital_Status_ID == "Single"), "Single", "Unknown")))))) %>%
            mutate(LatestSpell = ifelse(LatestNumberOfSpells < 5, "2 - 4 spells", "More than 4 spells"))


ad$BeenInpatientin12months[!is.na(ad$BeenInpatientin12months)] <- "Was an Inpatient"
ad$BeenInpatientin12months[is.na(ad$BeenInpatientin12months)] <- "Was Not an Inpatient"

ad$Gender_ID[ad$Gender_ID == "Not Specified"] <- NA
ad.use.gender.diff <- table(ad$PatientonAD, ad$Gender_ID)
ad.use.gender.diff <- ad.use.gender.diff[ ,-3]

ad.use.ethnicity <- table(ad$ethnicity, ad$PatientonAD)
ad.use.ethnicity <- ad.use.ethnicity[-36, ]
```

******
******

###Part 2: COMPARING ANTIDEPRESSANT USERS TO NON ANTIDEPRESSANT USERS


**Demographic Differences** 

```{r Chunk F3, echo = FALSE, results = "hide"}
x <- (CrossTable(x = ad$Gender_ID, y = ad$PatientonAD, prop.chisq = FALSE, prop.t = FALSE, prop.r = FALSE, prop.c = TRUE))
```

```{r Chunk F3.1, echo = FALSE}

kable((x$prop.tbl)*4157, 
      col.names = c("Not on AD", "On AD"), 
      digits = 0, 
      align = "c",
      caption = "Antidepressant Use Status by Gender (Number of Patients)")


kable((x$prop.row)*100, 
      col.names = c("Not on AD (%)", "On AD (%)"), 
      digits = 2, 
      align = "c",
      caption = "Antidepressant Use Status by Gender (row percentage)")
```

```{r Chunk F3.2, echo = FALSE, message=FALSE}
chisq.test(ad.use.gender.diff) 
```

******

_No difference in AD users and non-users by gender._  

******
******

```{r Chunk F4, echo=FALSE, results = "hide"}
z <- CrossTable(x = ad$ethnicity, y = ad$PatientonAD, prop.chisq = FALSE, prop.t = TRUE, prop.r = TRUE, prop.c = TRUE)
```

```{r Chunk F4.1, echo = FALSE}
kable((z$prop.tbl)*4157, 
      col.names = c("Not on AD", "On AD"), 
      digits = 0, 
      align = "c",
      caption = "Antidepressant Use Status by Ethnicity (totals)")



kable((z$prop.row)*100, 
      col.names = c("Not on AD (%)", "On AD (%)"), 
      digits = 2, 
      align = "c",
      caption = "Antidepressant Use Status by Ethnicity (row percentage)")

```

```{r Chunk F4.2, echo=FALSE}
chisq.test(ad.use.ethnicity) 
#chisq.test(z$prop.row) 
#chisq.test(z$prop.col)
```

******

_There is a difference in the proportion of patients on and off antidepressants by etnnicity. A lot more patients of Asian ethnicity are on Antidepressants compared to othe ethnic groups. There is some association between ethnicity and antidepressant use._ 

******
******

```{r Chunk G, echo = FALSE, results = "hide"}
# histogram, boxplot, scale it and do the same plots, test for normailty
# test for normality to decide what test to do
# https://stat.ethz.ch/R-manual/R-devel/library/stats/html/shapiro.test.html
#shapiro.test(ad$age)

# if it was a normal distribution
with(ad, t.test(age[PatientonAD == 1], age[PatientonAD == 0]))
```


**Age Distribution by Patient Antidepressant Use Status**
```{r Chunk H, echo = FALSE, fig.width = 10, fig.height = 10, fig.align= 'center'}
#qplot(as.factor(PatientonAD), age, data = ad, geom = c("jitter","boxplot"), color=as.factor(PatientonAD))
ad_agegraph <- ad
ad_agegraph$PatientonAD[ad_agegraph$PatientonAD == 0] <- "No"
ad_agegraph$PatientonAD[ad_agegraph$PatientonAD == 1] <- "Yes"

qplot(as.factor(PatientonAD), age, data = ad_agegraph, 
      geom = c("jitter","boxplot"), 
      fill=as.factor(PatientonAD), 
      main = "Age Distribution by Patient Antidepressant Use Status") +
      theme(legend.position = "none") +
      xlab("Patient on Antidepressants?") +
      ylab("Age distribution")

#ad.use.age
ad_agegraph %>% select(age, PatientonAD) %>% 
            group_by(PatientonAD) %>% 
            summarise(mean(age),
                      sd = sd(age)) %>% 
  kable(digits = 2,
        align = "c",
        col.names = c("Patient on Antidepressants", "mean age (years)", "Standard Deviation"))

# if it isn't normal 
#http://www.gardenersown.co.uk/education/lectures/r/basics.htm#u_test
with(ad, wilcox.test(age[PatientonAD == 1], age[PatientonAD == 0]))
```

******

_Individuals taking antidepressants are older._

******
******

```{r Chunk I, echo = FALSE, results= "hide"}
ad.use.deprivation <- table(ad$imd_quartiles, ad$PatientonAD)

w <- CrossTable(x = ad$imd_quartiles, y = ad$PatientonAD, prop.chisq = FALSE, prop.t = TRUE, prop.r = TRUE, prop.c = TRUE)
```


```{r Chunk I1, echo = FALSE}
kable((w$prop.tbl)*4157, 
      col.names = c("Not on AD", "On AD"),
      digits = 0,
      align = "c",
      caption = "Antidepressant Use by (Residence) Area Deprivation (totals)")

kable((w$prop.row)*100, 
      col.names = c("Not on AD (%)", "On AD (%)"),
      digits = 2,
      align = "c",
      caption = "Antidepressant Use by (Residence) Area Deprivation (row percentages)")
```

_Row labels_: (1.64, 17.1] - Least deprived area of residence ; (48,63.5] - Most deprived area of residence

```{r Chunk I2, echo = FALSE}
chisq.test(ad.use.deprivation)
```

******

_The difference in deprivation quartiles by antidepressant use is significant. While there are similar proportions of individuals in each deprivation quartile on and off antidepressants, there is relatively less representation of patients from the extreme levels of deprivation._ 

******

**Clinical Differences** 

I used "Past Inpatient Status", "Number of prior Admissions to the Trust" and "HoNOS Total" to be proxy measures of _clinical severity_.


```{r Chunk H1, echo = FALSE, results = "hide"}
ad.use.spell <- table(ad$LatestSpell, ad$PatientonAD)

q <- CrossTable(x = ad$LatestSpell, y = ad$PatientonAD, prop.chisq = FALSE, prop.t = TRUE, prop.r = TRUE, prop.c = TRUE)
```

```{r Chunk H1.2, echo = FALSE, warning = "hide"}
qplot(LatestNumberOfSpells, data = ad, binwidth = 1, 
      color = I('white'), 
      fill = 'red',
      main = "Number of Admissions",
      xlab = "Spell Number (Admissions)")

kable((q$prop.row)*100, 
      col.names = c("Not on AD (%)", "On AD (%)"),
      digits = 2, 
      caption = "Number of Past Admissions (Spells) To Trust by Antidepressant Status")

chisq.test(ad.use.spell) 
```

******

_There is no association between number of admissions and antidepressant use._

******
******

```{r Chunk K, echo = FALSE, results = "hide"}
ad.use.inpatient <- table(ad$BeenInpatientin12months, ad$PatientonAD)

e <- CrossTable(x = ad$BeenInpatientin12months, y = ad$PatientonAD, prop.chisq = TRUE, prop.t = TRUE, prop.r = TRUE, prop.c = TRUE)
```

```{r Chunk K1, echo = FALSE}
kable((e$prop.row)*100, 
      col.names = c("Not on AD (%)", "On AD (%)"),
      digits = 2, 
      caption = "Past Inpatient Status by Antidepressant Status")

chisq.test(ad.use.inpatient)
```

******
******

_More individuals who were inpatients in the past, are currently on antidepressants._

******

```{r Chunk K2, echo = FALSE, results = "hide"}
ad.use.honos.total <- table(ad$Adjusted_Total_level, ad$PatientonAD)
#ad.use.honos.total <- ad.use.honos.total[-3, ]
f <- CrossTable(x = ad$Adjusted_Total_level, y = ad$PatientonAD, prop.chisq = FALSE, prop.t = TRUE, prop.r = TRUE, prop.c = TRUE)
```

```{r Chunk K3, echo = FALSE}
kable((f$prop.row)*100, 
      col.names = c("Not on AD (%)", "On AD (%)"),
      digits = 2, 
      caption = "Recent HoNOS total by Antidepressant Status")

chisq.test(ad.use.honos.total)
```

******

_Clinical severity doesn't seem to be associated with antidepressant use._

******
******

###Part 3: Antidepressant Use and Suicide Attempt

There is some evidence that antidepressant use may lead to suicidal ideation and attempts. Here i will explore the suicide attempt variable and ascertain whether it is associated / correlated with antidepressant use. 

If there is a significant association, can we pinpoint to a specific class of antidepressants?

******

```{r Chunk L1, echo = FALSE, results = "hide"}
ad$SuicideAttemptRecent[!is.na(ad$SuicideAttemptRecent)] <- "Yes"
ad$SuicideAttemptRecent[is.na(ad$SuicideAttemptRecent)] <- "No"


ad.use.suicide.attempt <- table(ad$SuicideAttemptRecent, ad$PatientonAD) 

CrossTable(x = ad$SuicideAttemptRecent, y = ad$PatientonAD, prop.chisq = FALSE, prop.t = TRUE, prop.r = TRUE, prop.c = TRUE)
```

```{r Chunk L2, echo = FALSE}
kable(ad.use.suicide.attempt,
      col.names = c("Not on Antidepressants", "On Antidepressants"))

chisq.test(ad.use.suicide.attempt)
```

******

_What do the above tables and test tell you about patients attempting suicide and antidepresant use?_

      - About **two thirds** of my cohort have attempted suicide within the snapshot period. 
      
      - There is a difference in the proportion of individuals who attempt suicide in 
      
      antidepressant user versus antidepressant non user groups. 
      
      - This difference is significant with a p-value < 0.005
      

******








