Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r}
rm(list=ls())
gc()

setwd("/Users/andreafernandes/Desktop/NewTestFolder/Lesson5_EDA/")
pf <- read.table("/Users/andreafernandes/Desktop/NewTestFolder/Lesson4_EDA/pseudo_facebook.tsv", sep="\t", header = TRUE)

library(dplyr)
```


```{r Third Qualitative Variable}
pf.fc_by_age_gender <- pf %>% 
                              filter(!is.na(gender)) %>%
                              group_by(age, gender) %>%
                              summarise(mean_friend_count =  mean(friend_count),
                              median_friend_count = median(friend_count),
                              n = n()) %>%
                              arrange(age)

by(pf.fc_by_age_gender$mean_friend_count, pf.fc_by_age_gender$gender, summary)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
library(ggplot2)

ggplot(aes(x = age, y = median_friend_count ), 
       data = pf.fc_by_age_gender) +
       geom_line(aes(color = gender))

```

***

### Thinking in Ratios
Notes:
```{r}
library(reshape2)
pf.fc_by_age_gender_wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'median_friend_count')

str(pf.fc_by_age_gender_wide)


ggplot(aes(x = age, y = male/female), 
       data = pf.fc_by_age_gender_wide) +
       geom_line(colour = 'green') +
       geom_hline(aes(yintercept=1), colour = 'blue', linetype = 2)

```
***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(x = age, y = male/female), 
       data = pf.fc_by_age_gender_wide) +
       geom_line(colour = 'green') +
       geom_hline(aes(yintercept=1), colour = 'blue', linetype = 2)

```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
str(pf)

#year_joined: Year that they joined facebook
#Current year is 2014
#tenure is duration as fb member


#my way 
#pf <- pf %>% select(tenure) %>% 
#       mutate (Curryear = as.Date("01/06/2014", format = "%d/%m/%y")-(365.25*5.998)) %>%
#       mutate (year_joined = year(Curryear) - floor((tenure/365.25))) %>% tail()

#another (more easier) way to do this. 

pf$year_joined <- floor(2014 - pf$tenure/365)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
pf <- pf %>% mutate(year_joined.bucket= cut(year_joined, 
                                          breaks=c(2004, 2009, 2011, 2012, 2014))) 

```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count),
           data = subset(pf, !is.na(pf$year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
           data = subset(pf, !is.na(pf$year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median) +
  geom_line(stat = 'summary', fun.y =  mean, linetype = 2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
str(pf)

pf$friend_rate <- pf$friend_count/pf$tenure


with(subset(pf, tenure >=1, !is.na(tenure)), summary(friend_count/tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate? .2205

What is the maximum friend rate? 417

```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated/tenure),
           data = na.omit(pf), subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket), 
            stat = 'summary', 
            fun.y = mean) 
#+
#  coord_cartesian(xlim = c(1400, 1600), ylim = c(0, 1.5))
```

***

### Bias-Variance Tradeoff Revisited
Notes: The code changing the binning is substituting x = tenure in the plotting expressions with x = 7 * round(tenure / 7), etc., binning values by the denominator in the round function and then transforming back to the natural scale with the constant in front.

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***
