---
title: "Corruption"
author: "andreafernandes"
date: "3 April 2016"
output: html_document
---



```{r Chunk A, echo=TRUE, results="hide"}
rm(list=ls())
gc()


suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tidyr))

setwd("/Users/andreafernandes/Desktop/Google Drive/Springboard_Data_Science_Course_2016/DataScience/Lesson3/4_CorruptionDataset/")

```

******

Loading the datasets
```{r Chunk A1, echo=TRUE, results="hide"}
Corruption <- read.csv("/Users/andreafernandes/Desktop/Google Drive/Springboard_Data_Science_Course_2016/DataScience/Lesson3/4_CorruptionDataset/2015_CPI_data.csv")

Corruption_past <- read.csv("/Users/andreafernandes/Desktop/Google Drive/Springboard_Data_Science_Course_2016/DataScience/Lesson3/4_CorruptionDataset/Corruption.csv")
```

******

What do the datasets look like? 

```{r Chunk B}
data.frame(head(Corruption))

data.frame(head(Corruption_past))

```

Things to do

        1) Rename variables
        2) Convert cpi in cp dataset (cpi.in.08 and cpi.in.09) to numbers comparable to cpi.in.15 (multiply by 10)
        3) Select only certain variables from Corruption 2015 dataset and 
        4) Merge datasets together
      
      
***

```{r Chunk B1}
#Selecting key variables from Corruption 2015 dataset

c <- Corruption %>% 
          select(Country, Rank, cpi.in.15 = CPI2015, Region, World.Justice.Project.ROL, Number.of.Sources)



# Renaming variables and making cpi figures comparable to 2015
cp <- Corruption_past %>% rename(Country = X, cpi.in.08 = X2008, cpi.in.09 = X2009) %>%
                          mutate(cpi.in.08 = cpi.in.08*10) %>% 
                          mutate(cpi.in.09 = cpi.in.09*10)



#full join datasets
c.merge <- inner_join(c, cp)

data.frame(head(c.merge))

c.merge <- c.merge %>% select(Country, Region, cpi.in.08, cpi.in.09, cpi.in.15, World.Justice.Project.ROL, Number.of.Sources)
```


*******

A little bit about the dataset: 

CPI stands for Corruption Perception Index. 

cpi.in.08, cpi.in.09 and cpi.in15 are all CPI scores in the year 2008, 2009 and 2015, respectively. The higher the **Score** the less the corruption

_**Questions to answer to help with EDA of this dataset.** 

The aim is to try and get to "know" this dataset to make an informed hypothesis for data analysis. 

      1) What country has the highest corruption perception in 2008?
      2) What country has the highest corruption preception in 2009?
      3) Which country did the perception of corruption increase the most?
      4) What is the World Justice Project?
      5) Could you use it to tell you more about the corruption perceptions?
      6) Does the number of sources affect the corruption perception you think?
      7) What about countries like India where the corruption levels are high, how do they compare to the corruption perceptions the rest of the world?
      
```{r Chunk C}
#my solution
c.merge %>% arrange(cpi.in.08) %>% slice(1:5) 

c.merge %>% arrange(cpi.in.09) %>% slice(1:5) 

c.merge %>% arrange(cpi.in.15) %>% slice(1:5) 

#much easier solution
subset(c.merge, c.merge$cpi.in.08 == min(c.merge$cpi.in.08, na.rm = TRUE)) #Somalia

subset(c.merge, c.merge$cpi.in.09 == min(c.merge$cpi.in.09, na.rm = TRUE)) #Somalia

subset(c.merge, c.merge$cpi.in.15 == min(c.merge$cpi.in.15, na.rm = TRUE)) #Somalia

```


What's corruption perception like in the three years?
```{r Chunk D}
c.m <- gather(c.merge, Year, Cpi, cpi.in.08:cpi.in.15)

ggplot(aes(x = Country, y = Cpi, color = Country), data = c.m) +
  geom_bar(stat = "identity", width = 0.5) +
  facet_grid(Year ~ Region, scales ="free_x") + 
      theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1))
```

Summary:

        i) In 2008, Europe has one of the highest scores for low corruption perception; with the African and East and Central african regions having some of the lowest scores for low corruption perception. 
        ii) The trends of CPI across the years seem to be unchanged. 
          
          
The variation in CPI over the years...
```{r Chunk E}
qplot(x = Year, y = Cpi, data = c.m, geom = 'boxplot', color = I('white'), fill = I('#FF0000')) +
coord_flip()

```

Summary
      
        i) In 2015, compared to 2008 and 2009 (5/6 years ago), the corruption perception improves a little bit. 


The variation in CPI over the years by region...
```{r Chunk E1}
qplot(x = Year, y = Cpi, data = c.m, geom = 'boxplot', color = I('white'), fill = I('#FF0000')) +
      coord_flip() +
      facet_wrap(~Region)
```
  
Summary:

      i) Apart from Middle east and Europe (where corruption perception is positive anyway), Americas, Asia and Africa show improvement in corruption scores over the year.  
```{r Chunk F}
head(data.frame(c.merge))

qplot(x = cpi.in.15, y = World.Justice.Project.ROL, data = c.merge) 
  

cor.test(x=c.merge$cpi.in.15, y = c.merge$World.Justice.Project.ROL, method = "pearson", conf.level = 0.95)
```

Summary

      i) There is a strong correlation with Justice scores and high no corruption perception levels. 


```{r Chunk G}
table(c.merge$Number.of.Sources)

qplot(x = cpi.in.15, data = c.merge) + facet_wrap(~Number.of.Sources)

ggplot(aes(x = Country, y = cpi.in.15), data = c.merge) +
  geom_bar(stat = "identity", width = 0.5) +
  facet_wrap(~Number.of.Sources, scales ="free_x") + 
      theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1))

```

Summary:

      i) Most countries have 7 to 8 sources to provide an overall corruption perception index. 
      ii) Countries that require less sources or more than 8 sources are developing eastern European, African and Middle Eastern countries


```{r Chunk G1}
qplot(x = cpi.in.15, y = World.Justice.Project.ROL, data = c.merge) +
  facet_wrap(~Region)
```

Summary:

      i) The positive correlation between Justice score and corruption index is reflected in all regions of the world. 
      
```{r Chunk G2}
qplot(x = cpi.in.15, y = World.Justice.Project.ROL, data = c.merge, color = Country) +
  facet_grid(Region ~ Number.of.Sources, scales ="free_x") +
  theme(legend.position = "none")

```


```{r Chunk H}
c.merge <- c.merge %>% mutate(diff = cpi.in.15 - cpi.in.08)

ggplot(aes(x = Country, y = diff, colour = I('white'), fill = I('#FF0000')), data = c.merge) +
  geom_bar(stat = "identity", width = 0.5) +
  facet_wrap(~Region, scales ="free_x") + 
      theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1))

```

Summary:
  
      i) Apart from the China outlier (this is a poor naming of Macau and mainland China), Oman seems to be the country with worsening corruption perception index...
      ii)...While Rwanda has the most improvement in corruption perception index. See below
      
      
```{r Chunk H2}
subset(c.merge, c.merge$diff == min(c.merge$diff, na.rm = TRUE)) 

subset(c.merge, c.merge$diff == max(c.merge$diff, na.rm = TRUE)) 

c.merge %>% arrange(diff) %>% slice(1:5) 
```





```{r Chunk I, Miscellaneous code, echo = FALSE, results ="hide"}
qplot(x=Country, y = CPI2015, data = Corruption, na.rm = TRUE) +
      facet_wrap(~Region, scales = "free_x") +
      geom_hline(aes(yintercept=0, colour = 'red')) + 
      theme(axis.text.x = element_text(angle = 60, hjust = 1))

```