---
title: "Yoghurt"
author: "andreafernandes"
date: "1 April 2016"
output: html_document
---

### Introducing the Yogurt Data Set
Notes:
```{r Chunk A}
rm(list=ls())
gc()

setwd("/Users/andreafernandes/Desktop/NewTestFolder/Lesson5_EDA/Lesson5_yoghurt_EDA/")

yo <- read.csv("/Users/andreafernandes/Desktop/NewTestFolder/Lesson5_EDA/Lesson5_yoghurt_EDA/yogurt.csv")

str(yo)

yo$id <- factor(yo$id)
```
***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
library(ggplot2)

qplot(x = price, data = yo, binwidth = 1) +
      scale_x_continuous(lim = c(10, 70), breaks = seq(10, 70, 10))


summary(yo$price)
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
unique(yo$price)

length(unique(yo$price))


library(tidyr)

head(data.frame(yo))

yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)


head(data.frame(yo))

```

***

### Prices over Time
Notes:

```{r Prices over Time}

ggplot(aes(x = time, y = price), data = yo) + 
  geom_smooth() +
  geom_smooth(stat = 'summary', fun.y = median) + 
  scale_x_sqrt() 
  #coord_cartesian(xlim = c(20, 50), ylim = c(9600, 9700))
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)

sample.ids <- sample(levels(yo$id), 16)




set.seed(1234)

sample.myIDs <- sample(levels(yo$id), 20)

ggplot(aes(x = time, y = price),
       data = subset(yo, id %in% sample.myIDs)) + #create a plot for every id in sample.myIDs
       facet_wrap( ~ id) +
       geom_line() + 
       geom_point(aes(size = all.purchases), pch = 1)

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r}
library(GGally)

```

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

