---
title: "Diamonds_Part2"
author: "andreafernandes"
date: "29 March 2016"
output: html_document
---


###Diamonds Dataset. Project Part 1

####Using the diamonds dataset to practice a few EDA techniques and tools. 

Working through the exercises in Problem Set 4: Exploratory Data Analysis with R


```{r, Chunk A, echo=FALSE, results = "hide"}
rm(list=ls())
gc()

suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(gridExtra))

setwd("/Users/andreafernandes/Desktop/Google Drive/Springboard_Data_Science_Course_2016/DataScience/Lesson4_EDA/1_Diamonds_Part2/")
```


******



a) Load the diamonds data set
```{r Chunk A}
head(data.frame(diamonds))
```


******

Your first task is to create a scatterplot of price vs x. using the ggplot syntax.
```{r Chunk B}
qplot(x = x, y = price, data = diamonds) +
      geom_point(alpha = 1/20) 

summary(diamonds$price)
summary(diamonds$x)
subset(diamonds, diamonds$x == min(diamonds$x, na.rm = TRUE))

diamonds %>% arrange(x) %>% slice(1:15)
```

Summary of the scatterplot

******

What is the correlaton between price and x, y and z?
```{r Chunk C}
cor.test(x = diamonds$price, y =  diamonds$x, method = "pearson", conf.level = 0.95)

cor.test(x = diamonds$price, y =  diamonds$y, method = "pearson", conf.level = 0.95)

cor.test(x = diamonds$price, y =  diamonds$z, method = "pearson", conf.level = 0.95)
```

******

Create a simple scatter plot of price vs depth.
```{r Chunk D}
qplot(y = price, x = depth, data = diamonds)


# Change the code below to make the transparency of the
# points to be 1/100 of what they are now and mark
# the x-axis every 2 units. See the instructor notes
# for two hints.

ggplot(data = diamonds, aes(x = depth, y = price)) + 
  geom_point(alpha = 1/100) + 
  scale_x_continuous(lim = c(0, 80), breaks = seq(0,80, 2))
```


******


What is the correlation of diamond price with depth?
```{r Chunk E}
cor.test(x = diamonds$price, y =  diamonds$depth, method = "pearson", conf.level = 0.95)
```

*****

Create a scatterplot of price vs carat and omit the top 1% of price and carat values.
```{r Chunk F}
ggplot(data = diamonds, aes(x = carat, y = price)) +
  geom_point()

ggplot(data = diamonds, aes(x = carat, y = price)) + 
  geom_point(alpha = 1/10, stat = 'summary', fun.y = quantile, fun.args=list(probs = 0.99)) 
```


Create a scatterplot of price vs. volume (x * y * z).
This is a very rough approximation for a diamond's volume.

Create a new variable for volume in the diamonds data frame.

Don't make any adjustments to the plot just yet.

```{r Chunk G}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z

ggplot(data = diamonds, aes(x = volume, y = price)) +
  geom_point()
```

Summary:

The average volume of the diamonds is below 800. 

There is an outlier at around less than 4000. 

The higher the volume the higher the prices. 

******


What's the correlation of price and volume? Exclude diamonds that have a volume
of 0 or greater than or equal to 800. 
```{r Chunk H}
d <- filter(diamonds, !diamonds$volume == 0, !diamonds$volume >= 800)

cor.test(x = d$price, y =  d$volume, method = "pearson", conf.level = 0.95)
```


******

Subset the data to exclude diamonds with a volume greater than or equal to 800. Also, exclude diamonds
with a volume of 0. Adjust the transparency of the points and add a linear model to the plot. (See the
Instructor Notes or look up the documentation of geom_smooth() for more details about smoothers.)

```{r Chunk I}
ggplot(aes(x = volume, y = price), data = subset(diamonds, diamonds$volume > 0 & diamonds$volume < 800)) +
  geom_jitter(alpha = 1/5) + 
  geom_smooth()
```


*******

Do you think this would be a useful model to estimate the price of diamonds? Why or why not?

The model predicts the data up until volume 400. After that, it incorrectly models the data, this could be due to the increased variance in price after volume 400. 

(https://discussions.udacity.com/t/lesson-4-diamonds-volume-by-price/163306)

*******


Use the function dplyr package to create a new data frame containing info on diamonds by clarity.

Name the data frame diamondsByClarity.

The data frame should contain the following variables in this order.

     (1) mean_price
     (2) median_price
     (3) min_price
     (4) max_price
     (5) n
where n is the number of diamonds in each level of clarity.
```{r Chunk J}
diamondsByClarity <- diamonds %>% group_by(clarity) %>%
             summarise(mean_price = mean(price),
                       median_price =  median(price),
                       min_price = min(price), 
                       max_price = max(price), 
                       n = n())

```

******


We’ve created summary data frames with the mean price by clarity and color. You can run the code in R to
verify what data is in the variables diamonds_mp_by_clarity and diamonds_mp_by_color.

```{r Chunk K}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))
```

Your task is to write additional code to create two bar plots on one output image using the grid.arrange() function from the package gridExtra.
```{r Chunk L}
p1 <- ggplot(aes(x = clarity, y = mean_price, colour = I('white'), fill = I('#FF0000')), data = diamonds_mp_by_clarity) + 
      geom_bar(stat = "identity")

p2 <- ggplot(aes(x = color, y = mean_price, colour = I('white'), fill = I('#FF0000')), data = diamonds_mp_by_color) + 
      geom_bar(stat = "identity")

grid.arrange(p1, p2)

```
Summary:
        The lowest prices are for the "best" color D and E. 
        Whereas the best clarity I1 and SI2 cost more than relatively lower clarity diamonds. 
        Clarity trumps color in terms of price it seems. (see below though, it doesn't seem to be the case)
        
        
```{r Chunk M}

qplot(x = color, y = price/1000, data = diamonds) + 
      geom_bar(stat = "identity", colour = I('white'), fill = I('#FF0000')) +
      facet_wrap(~clarity, scales =  "free_y") 

```

Summary: 


There are a whole host of factors that affect diamond price. The cut, clarity, color, size, weight and demand for the diamonds. 

(http://www.lumeradiamonds.com/diamond-education/index)